Managing Traceability Links between Specification and Implementation of Hardware/Software Interfaces

### 2013-05-05 RPE/implementation.tex
revise

This section presents the architecture of the implementation of our approach first. After that, we introduce the algorithm for validating the traceability links, and for code selection completion which helps the user to complete the code piece selection when part of a syntax element is selected.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.55\textwidth]{architecture}
\caption{coDoc Architecture and Features}
\label{fig:architecture}
\end{center}
\end{figure}

The architecture of our approach is shown in Figure \ref{fig:architecture}. 
There are three types of data here, 1) code, 2) spec, and 3) traceability links.
These data are consumed by three components: code manager (CM), spec manager (SM) and link manager (LM).
\begin{enumerate}
\item \textit{Link Manager.} Traceability link manager get the highlighted code piece from code manager and the highlighted spec piece from spec manager, and create traceability link for these two piece of information. It also reads the traceability link data, verify the validity of the link,
and highlights the code and document pieces that are related.
\item \textit{Code Manager.} Code manager reuses C/C++ Development Tool (CDT) editor from Elipse\footnote{\texttt{\url{http://www.eclipse.org}}} to parses the code, 
and highlights syntax elements. It can highlight the piece of code based on the information provided by the user or the link manager.
\item \textit{Spec Manager.} Spec manager renders the pdf content, can support select and highlight spec content, and communicate with the link manager.
\end{enumerate}

We implemented coDoc as an Elipse Rich Client Program (RCP), 
which can make our tool share the appearence in Eclipse that is familiar by many programmer.
Figure \ref{fig:platformview} is the screenshot of the tool.
The bottom is the UI for link manager, while the top left part and top right side are the UI for code manager and spec manager repectively.

% convert platformview.png platformview.eps
\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{platformview}
\caption{coDoc Platform View}
\label{fig:platformview}
\end{center}
\end{figure}

@@

\begin{algorithm}[!h]
\DontPrintSemicolon
\SetKwComment{tcp}{\hfill$\triangleright$ }{}
\SetCommentSty{emph}
$AST \leftarrow \textsc{Gen-AST}(code)$\;
$\langle RES, CONTENT_C \rangle \leftarrow \textsc{Get-Code-Content}(AST, link.code.path)$\;
\If{$RES = FAIL$} {
\Return $FAIL$\;
}
$RES \leftarrow \textsc{Compare-Content}(CONTENT_C, link.code.content)$\;
\If{$RES = FAIL$} {
\Return $FAIL$\;
}
$CONTENT_S \leftarrow \textsc{Get-Spec-Content}(spec, link.spec.path)$\;
$RES \leftarrow \textsc{Compare-Content}(CONTENT_S, link.spec.content)$\;
\If{$RES = FAIL$} {
\Return $FAIL$\;
}
\Return $SUCCESS$\;
\caption{$\textsc{Traceability-Link-Validation}(link)$}
\label{alg:validate}
\end{algorithm}

@@

When code evolves, we need to validate the existing traceability links.
Once an invalid traceability links is found, we should report it to let user correct it manually.
To check the validity of the traceability links, we record the content of the code piece and the spec piece when generating the links.
The content is used later to compare with the content to which the path points to when code changes.
The validation process basically go through each traceability link in the database, 
and check whether the link is valid or not.
Algorithm~\ref{alg:validate} presents the pseudo-code for check the validation of one single traceability link.
Line 1 - line 7 check the validity of the code piece part of the traceability link.
Line 1 generates the AST for the current code first.
Line 2 then invokes the function \textsc{Get-Code-Content} with the path of the code piece as input, identify the code piece in the new AST, and return the content of the code piece if the code piece is identified successfully.
If the path of the code piece cannot identify any code piece or can identify more than one code piece, the function \textsc{Get-Code-Content} reports failure.
Line 5 compare the the new content extracted from the new code with the recorded old content, and fail if the two contents do not match.
After finishing check the code piece part, line 8 - line 11 of the algorithm check the spec piece part in a similar flow.
If no failure happens, the algorithm return successfully at line 12.

@@

\begin{algorithm}[!h]
\DontPrintSemicolon
\SetKwComment{tcp}{\hfill$\triangleright$ }{}
\SetCommentSty{emph}
$AST \leftarrow \textsc{Gen-AST}(code)$\;
$SEQUENCE_{LEAF} \leftarrow \textsc{Gen-Leaf-Sequence}(AST)$\;
$POC_{AST} \leftarrow \textsc{To-AST-Based}(SEQUENCE_{LEAF}, selection_{org})$\;
$selection_{new} \leftarrow \textsc{To-Offset-Based}(POC_{AST})$\;
\Return $selection_{new}$\;
\caption{$\textsc{Complete-Code-Selection}(selection_{org})$}
\label{alg:completion}
\end{algorithm}

To improve the user experience of coDoc, we help users to select the code pieces when creating traceability links.
Our approach is based on the fact that it is usually meaningless to have part of the basic code elements as the code piece of the traceability links.
For example, we do not create a link between half of the variable and the spec piece.
We assume that every selected piece of code is composed of valid syntax elements.
Based on this assumption, we can help the user to automatically complete their selection when they only select part of the syntax element, like a variable, a constant, and etc. 
Algorithm \ref{alg:completion} outlines our approach.
This algorithm takes a code selection in offset based format,
and returns a new code selection in offset based format.
We choose offset based format because Eclipse use this format to highlight the code piece.
The algorithm basically includes two parts: (1) line 1 - line 3 force the input selection to be aligned by transforming it from offset based code piece to AST based code piece. (2) line 4 generates the required format by transforming AST based code piece back to offset based code piece.
As we discussed before, AST based representatioin is a sequence of AST leaf nodes in DFS visit order, which is the same order as in the source code.
The AST based code piece covers all the meaningful part of the original offset based code piece.
To transform the offset based code piece to AST based code piece, 
we get the sequence of leaf nodes of AST in line 1 and line 2,
and then match the leaf node sequence with the code piece to get the AST based representation.
The leaf node in the AST based representation is determined this way: if the leaf node contains any part of the selected code, then the leaf node is in the representation.
After we get the AST based representation,
we transform it back to offset based reprensentation, and return it.
This way, each time the user make a selection,
our engine adjust the selection to valid syntax boundary.

>>>>

This section first presents the architecture of the coDoc prototype tool, which implements our traceability link management framework. After that, we introduce the algorithms for validating the traceability links and for code selection completion which helps users to complete code piece selection when part of a syntax element is selected.

@@

The architecture of our approach is shown in Figure \ref{fig:architecture}.
There are three types of data here, 1) code, 2) spec, and 3) traceability links.
These data are consumed by three components: code manager (CM), spec manager (SM) and link manager (LM).
\begin{figure}
\begin{center}
\includegraphics[width=0.55\textwidth]{architecture}
\caption{coDoc Architecture and Features}
\label{fig:architecture}
\end{center}
\end{figure}
\begin{enumerate}
\item \textit{Link Manager.} Traceability link manager gets the highlighted code piece from code manager and the highlighted spec piece from spec manager, and creates traceability link for these two piece of information. It also reads the traceability link data, verifies the validity of the link,
and highlights the code and document pieces that are related.
\item \textit{Code Manager.} Code manager reuses C/C++ Development Tool (CDT) editor from Eclipse~\cite{Eclipse} to parses the code,
and highlights syntax elements. It can highlight the piece of code based on the information provided by the user or the link manager.
\item \textit{Spec Manager.} Spec manager renders the pdf content, can support select and highlight spec content, and communicate with the link manager.
\end{enumerate}

We implemented coDoc as an Elipse Rich Client Program (RCP),
which can make our tool share the appearance in Eclipse that is familiar by many programmer.
Figure \ref{fig:platformview} is the screenshot of the tool.
The bottom is the UI for link manager, while the top middle part and top right side are the UI for code manager and spec manager respectively.
To traverse a traceability link, users can select a link from the list in the link manager or click on a code construct in the code manager. 
As a result, the link, the relevant code, and the relevant spec piece are highlighted in the link, code, and the spec mangers.

% convert platformview.png platformview.eps
\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{platformview}
\caption{coDoc User Interface Screenshot}
\label{fig:platformview}
\end{center}
\end{figure}

@@

@@

When code evolves, we need to validate the existing traceability links.
Once an invalid traceability link is found, we report it and ask the user to correct it manually.
To check the validity of the traceability links, we record the content of the code piece and the spec piece when generating the links.
The content is used later to compare with the content to which the path points to when code changes.
The validation process basically goes through each traceability link in the database,
and checks whether the link is valid or not.
\begin{algorithm}[!h]
\DontPrintSemicolon
\SetKwComment{tcp}{\hfill$\triangleright$ }{}
\SetCommentSty{emph}
$AST \leftarrow \textsc{Gen-AST}(code)$\;
$\langle RES, CONTENT_C \rangle \leftarrow \textsc{Get-Code-Content}(AST, link.code.path)$\;
\If{$RES = FALSE$} {
\Return $FAILURE$\;
}
$RES \leftarrow \textsc{Compare-Content}(CONTENT_C, link.code.content)$\;
\If{$RES = FAIL$} {
\Return $FAILURE$\;
}
$CONTENT_S \leftarrow \textsc{Get-Spec-Content}(spec, link.spec.path)$\;
$RES \leftarrow \textsc{Compare-Content}(CONTENT_S, link.spec.content)$\;
\If{$RES = FAIL$} {
\Return $FAILURE$\;
}
\Return $SUCCESS$\;
\caption{$\textsc{Traceability-Link-Validation}(link)$}
\label{alg:validate}
\end{algorithm}
Algorithm~\ref{alg:validate} presents the pseudo-code for check the validation of one single traceability link.
Line 1 - line 7 check the validity of the code piece part of the traceability link.
Line 1 generates the AST for the current code first.
Line 2 then invokes the function \textsc{Get-Code-Content} with the path of the code piece as input, identify the code piece in the new AST, and return the content of the code piece if the code piece is identified successfully.
If the path of the code piece cannot identify any code piece or can identify more than one code piece, the function \textsc{Get-Code-Content} reports failure.
Line 5 compares the the content extracted from the new code with the recorded content, and reports failure if they do not match.
After finishing checking the code piece, line 8 - line 11 of the algorithm check the spec piece in a similar flow.
If no failure happens, the algorithm returns successfully at line 12.

@@

%To improve the user experience of coDoc, we help users to select the code pieces when creating traceability links.
The coDoc tool assists user to select the code pieces that can be used in traceability links. Each code piece selected is required to be a complete language construct.
%Our approach is based on the fact that it is usually meaningless to have part of the basic code elements as the code piece of the traceability links.
For example, we do not create a link between part of a variable name in the code and part of a word in the spec.
We assume that every selected piece of code is composed of valid syntax elements.
Based on this assumption, we can help users to automatically complete their selection when they only select part of a syntax element such as a variable. 
Algorithm \ref{alg:completion} outlines our approach.
\begin{algorithm}[!h]
\DontPrintSemicolon
\SetKwComment{tcp}{\hfill$\triangleright$ }{}
\SetCommentSty{emph}
$AST \leftarrow \textsc{Gen-AST}(code)$\;
$LEAF\_SEQ \leftarrow \textsc{Gen-Leaf-Sequence}(AST)$\;
$AST\_selection \leftarrow \textsc{To-AST-Based}(LEAF\_SEQ, original\_offset)$\;
$new\_offset \leftarrow \textsc{To-Offset-Based}(AST\_selection)$\;
\Return $new\_offset$\;
\caption{$\textsc{Complete-Code-Selection}(orginal\_offset)$}
\label{alg:completion}
\end{algorithm}
This algorithm takes a code selection in the offset based format,
and returns a new, appropriately enlarged code selection in the offset based format.
We choose offset based format because Eclipse use this format to highlight code pieces.
The algorithm basically works as follows: (1) lines 1 - 3 align the input offset-based selection to its smallest enclosing language construct by analyzing the AST.
(2) line 4 maps the enlarged code selection to the offset based format and returns it.  
As we discussed above, the AST based selection is a sequence of AST leaf nodes in the depth-first-search order, which is the same order as the code in the source code.
The AST based code piece covers all the meaningful part of the original offset based code piece.
To transform the offset based selection to the AST based selection,
we obtain the sequence of leaf nodes of the AST in line 1 and line 2,
and then match the leaf node sequence with the offset-based selection to get the enlarged AST based selection.
The leaf nodes in the AST based selection is determined as follows: if a leaf node contains any part of the offset selected code, then the leaf node is in the AST based selection.
After we obtain the AST based selection, we transform it back to a new, enlarged offset based selection, and display it to users.
Each time users make a selection, coDoc adjusts the selection to valid syntax boundary.

### 2013-05-05 RPE/introduction.tex
revise

The majority of software engineering work is devoted to maintaining legacy software systems other than developing new ones \cite{abran_maintenance_1991}.
It is estimated that up to 60\% of software maintenance time is spent on program comprehension \cite{lawrence_software_2006}.
Software developers often find it difficult to extract the knowledge about a software system,
because the knowledge is often implicitly expressed,
and distributed in a variety of software artifacts.
Examples of software artifacts include source code, design documents, application programming interface specifications, bug reports, etc.

To make the knowledge about a software system explicit,
reserachers have built various techniques to recover the semantic connections among software artifacts,
especially between documentation and source code \cite{antoniol_recoveringtraceability_2002} \cite{deLucia_incremental_2006} \cite{marcus_recovering_2003}.
These semantic connections are often called \textit{traceability links}. 
Effectively maintaining the traceability links can help improving the productivity of the developer and the quality of the software \cite{neumuller_case_2006}.

Unfortunately, these techniques are not ready to be applied to implementations and specifications of HW/SW interfaces,
even though these implementations are an important category of software that caused most of the system failures in both Windows and Linux operating systems \cite{swift_os_2003} \cite{chou_os_2001},
HW/SW interface forms the interaction point of hardware and software,
and defines the protocol for the hardware and software to communicate.
The implementation of HW/SW interfaces refers to hardware design, device drivers and virtual device prototypes, who implement the communication logic on both sides of the interfaces.
%The implements are designed for a variety of computer components, and form a large part of operating systems and virtual machines.
HW/SW interface specifications describe the HW/SW interfaces in natuaral lagnguage,
and guide the implementation of the interfaces.
HW/SW interface specifications play a crucial role in the comprehension and maintenance of the implementations of the interfaces.
The reason why existing traceability link recovering techniques do not work for HW/SW interfaces is that these traceability links need to be fine-grained,
which is not supported by the existing techniques.
HW/SW interfaces are mainly composed of software-programmable hardware registers mapped to the processor's address space.
The low-level abstraction of the software/hardware interface makes the traceability links fine-grained.
Traceability links between the specifications and implementations usually connect statements, which operate the registers, or expressions, which represent certain bits of registers, to the register bit descriptions in the specifications. 
However, previous tools usually maintain tracebility links between the classes or functions of the source code and sections or paragraphes of the documentation, which is more coarse-grained.

@@

We present a framework for managing traceability links between specifications and implementations of HW/SW interfaces. 
This framework supports fine-grained links between basic elements of specifications and constructs of implementation languages.
We design a data model to support the framework, and organize the data for the traceability links.
Our framework also supports automatic migration of links as implementations evolve.

We have implemented a prototype tool, namely coDoc, in support of this framework. 
With this tool, the developer can import and manage implementation source code and interface specifications, select a certain piece of implementation (POI) and a certain piece of specification (POS) of interest, create a traceability link connecting the seleted POI and POS, check out a different version of source code, review and verify the existing traceability links. 
We applied coDoc to three virtual device prototypes and their specifications, recovered traceability links, and verified the traceability links on other versions of the driver code.
Experimental results show that our framework fully supports the types of traceability links between the virtual devices and their specifications,
maintain traceability links with required granularity, 
and it accurately migrates valid links from version to version.
\todo{the data need to be updated}\hlc{It preserves 70\% of the existing traceability links when code evolves to a new version.}

This paper makes the following key contributions:
\begin{itemize}
\item Presented a framework for managing traceability links between specifications and implementations of HW/SW interfaces with accuracy, robustness and fine-granularity.
\item Developed coDoc, a tool implementing the approach.
\item Evaluated our approach with realistic examples.
\end{itemize}

The rest of the paper is organized as follows. 
Section \ref{sec:problem} examplifies and defines the problem that we address in this paper.
Section \ref{sec:approach} and section \ref{sec:implementation} illustrate the design and implementation details of our approach.
Section \ref{sec:evaluation} provides some preliminary evaluation results.
Section \ref{sec:related} reviews the related work.
We conclude our approach and discuss the future work in Section \ref{sec:conclusion}.

>>>>

Significant portion of software engineering efforts is devoted to maintaining legacy software systems~\cite{abran_maintenance_1991}.
It is estimated that up to 60\% of software maintenance efforts is spent on program comprehension~\cite{lawrence_software_2006}.
Software developers often find it difficult to extract the knowledge about a software system,
because such knowledge is often implicitly expressed and distributed in a variety of software artifacts.
Examples of software artifacts include source code, design documents, application programming interface specifications, bug reports, etc.

To capture such knowledge explicitly, various techniques have been developed to recover the semantic connections among software artifacts,
especially between documentation and source code \cite{antoniol_recoveringtraceability_2002} \cite{deLucia_incremental_2006} \cite{marcus_recovering_2003}.
Such semantic connections are often called \textit{traceability links}.
Effectively maintaining the traceability links helps improving developer productivity and software quality \cite{neumuller_case_2006}.

Unfortunately, these techniques are not readily applicable to specification and implementation of hardware/software (HW/SW) interfaces,
even though such implementation causes most system failures in both Windows and Linux operating systems \cite{swift_os_2003} \cite{chou_os_2001},
HW/SW interface inter-connects hardware and software and defines the protocol for hardware and software to communicate.
The implementation of HW/SW interfaces typically include hardware designs and device drivers. There is also software implementation of hardware design, e.g., virtual devices in virtual machines such as QEMU~\cite{Bellard05}
%and virtual device prototypes, who implement the communication logic on both sides of the interfaces.
%The implements are designed for a variety of computer components, and form a large part of operating systems and virtual machines.

A HW/SW interface specification describes a HW/SW interface in natural language and guides its implementation.
The specification plays a crucial role in the comprehension and maintenance of the implementation.
The key reason why existing traceability link recovering techniques do not work for HW/SW interfaces is that these traceability links need to be fine-grained,
which are not supported by the existing techniques.
HW/SW interfaces are mainly composed of software-programmable hardware registers mapped to the processor address space.
The low-level abstraction of the HW/SW interface demands the fine-grained traceability links.
These traceability links need to connect statements which operate the registers or expressions which represent certain bits of registers to the register bit descriptions in the specification. However, previous techniques maintain traceability links between classes or functions of the source code and sections or paragraphes of the specification, which is to coarse-grained.

@@

We present a framework for managing traceability links between specification and implementation of HW/SW interfaces.
This framework supports fine-grained links between basic elements of specifications and constructs of implementation languages.
A simple, but effective data model is developed to support the framework and organize the data of the traceability links.
Our framework also supports automatic migration of valid traceability links as implementation evolves.

We have implemented a prototype tool, namely coDoc, in support of this framework.
With coDoc, developers can (1) import and manage implementation source code and interface specifications, (2) select a piece of implementation (POI) and a piece of specification (POS) of interest, (3) create a traceability link connecting the selected POI and POS, (4) check out a different version of source code, and (5) migrate, review, and verify existing traceability links. We applied coDoc to source code of three virtual devices from QEMU and their specifications, recovered traceability links, and migrated these links to newer versions of source code .
Experimental results show that our framework fully supports the types of traceability links between the virtual devices and their specifications,
maintain traceability links with required granularity, and accurately migrates valid links from version to version.
\todo{the data need to be updated}\hlc{It preserves 70\% of valid traceability links when source code evolves to a new version.}

This paper makes the following key contributions:
\begin{itemize}
\item Presented a framework for managing traceability links between specification and implementation of HW/SW interfaces;  
\item Developed a simple, but effective data model for capturing and recording links; 
\item Identified appropriate anchors in implementation and specification that support fine-granularity and migration of valid links as implementation evolves; 
\item Developed coDoc, a prototype tool implementing the approach and evaluated our approach on realistic examples.
\end{itemize}

The rest of the paper is organized as follows.
Section \ref{sec:problem} elaborates on the problem addressed in this paper.
Section \ref{sec:approach} and Section \ref{sec:implementation} illustrate the design and implementation details of our approach.
Section \ref{sec:evaluation} provides preliminary evaluation results.
Section \ref{sec:related} reviews related work.
We conclude this paper and discuss future work in Section \ref{sec:conclusion}.

### 2013-05-05 RPE/paper.tex
revise

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage[pdftex]{graphicx}
\DeclareGraphicsExtensions{.pdf}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{float}
\usepackage{url}

@@

% first the title is needed
\title{Managing Traceability Links between Specifications and Implementations of Hardware/Software Interfaces}
% a short form should be given in case it is too long for the running head
\titlerunning{Traceability Links between Specifications and Implementations}
\authorrunning{Dejun Qian}

\author{Dejun Qian\\dejun@cs.pdx.edu}
\institute{Computer Science Department, Portland State University}

>>>>

\documentclass[letter]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage[pdftex]{graphicx}
\DeclareGraphicsExtensions{.pdf}
\usepackage{caption}
%\usepackage{subcaption}
\usepackage{float}
\usepackage{url}

@@

% first the title is needed
\title{Managing Traceability Links between Specification and Implementation of Hardware/Software Interfaces}
% a short form should be given in case it is too long for the running head
%\titlerunning{Traceability Links between Specifications and Implementations}
%\authorrunning{Dejun Qian}

\author{Dejun Qian}
\institute{Department of Computer Science  \\
Portland State University \\
Portland, OR 97207 \\
dejun@cs.pdx.edu}

### 2013-05-05 RPE/problem.tex
revise

% emphasize the special characteristic of hardware-related software
\section{Problem Statement}
\label{sec:problem}

\begin{figure}[h!]
  \begin{center}
    \begin{tabular}{c}
\begin{minipage}[b]{\linewidth}
  \centering
  \includegraphics[width=0.95\linewidth]{code1}
\end{minipage}\\
(a) The Old Version\\
\\
\begin{minipage}[b]{\linewidth}
  \centering
  \includegraphics[width=0.95\linewidth]{code2}
\end{minipage}\\
(b) The Newer Version
\end{tabular}
  \caption{Two Versions of a Virtual Devices Written in C. This piece of code is extracted from MAX7310 virtual device in QEMU virtual machine. This code defines a function to reset the registers. However, we do not know why they are resetted to the values labeled 1, 2 and 3. To know the answer, we need to refer to the specification.}
\label{fig:code}
\end{center}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.7\textwidth]{spec}
  \caption{Specification Example. This is part of MAX7310 specification. It includes three tables, which define the default values for three registers.}
  \label{fig:spec}
\end{figure}

The interfaces between software and hardware are composed of registers exposed by hardware devices and accessible by software. The implementation of the interfaces should conform to the specification in order to function correctly. Suppose we have to maintain and understand this implementation, we need to establish relationships between the implementation and the specification. Fig.~\ref{fig:code}(a) presents part of code extracted from a virtual device prototype for MAX7310 in QEMU. A virtual device prototype is a software implementation emulating a hardware device. MAX7310 is a 8-bit I/O port expander. QEMU (Quick EMUlator)\footnote{\texttt{\url{http://www.qemu.org}}} is virtual machine used widely in industry to assist hardware related software development. The source code for this implementation is written in C language. From the name of the function, we can guess the function shown is trying to reset the device to its initial state. However, we can not figure out what the values in the right sides of the assignment statements mean only by reading the code here. Fig.~\ref{fig:spec} includes the information extracted from the specification of MAX7310. In this figure, we can see three tables for three registers: polarity register, direction register and status register. These tables present the bit arrangement and the default value for each register. By looking at this specification, now we can see that the values used in the code are defined in the specification. The code uses the hex format, and the specification use binary format. The piece of code and the piece of specification labeled with the same number are related, which means the piece of specification explains the piece of code with the same label.

% in the directory including max7310.c: git log --follow --all -p max7310.c > a.txt
Code evolves all the time in its lifecycle. By checking the QEMU repository, we can find that the first version of the code in Fig~\ref{fig:code} was created on May 23, 2007, and the last change is on May 1, 2013. This code has changed 21 times till now, and it still keeps changing. 
Fig.~\ref{fig:code}(b) shows an example of a new version of the code in Fig.~\ref{fig:code}(a). Line 13, 25 and 29 of the code in Fig.~\ref{fig:code}(a) has changed to the line 12, 24 and 28 in Fig.~\ref{fig:code}(b).

Compared to code, specification is relatively stable. The specification in Fig~\ref{fig:spec} only has three versions, and the last change happened in 2005. The specification remains unchanged since the code was created.

This example is simple, but reveals three basic requirements for the management of traceability links between specification and implementations as follows.

\textbf{Requirement 1:} Traciability links should be able to uniquely refer to and accurately locate the relevant pieces of code in the source file and the piece of specification in the specification file. As in the example, there should be a way to locate the code pieces and specification pieces surrounded by the boxes labeled by 1, 2 and 3.

\textbf{Requirement 2:} Traceability links should be able to provide referrence at statement level or expression level for source code and at sentence level or word level for specification. As in this example, the code pieces and specification pieces surrounded by the boxes labeled by 1, 2 and 3 are all on this level.

\textbf{Requirement 3:} Traceability links should be immune from the evolusion of unrelated code. We should grarantee that the identifier created for the old version of code works for the new version of code if the code piece refered by the identifier does not change.

>>>>

% emphasize the special characteristic of hardware-related software
\section{Problem in Managing HW/SW Interface Traceability Links}
\label{sec:problem}

An interface between hardware and software is typically composed of registers exposed by hardware devices and accessible by software. The implementation of the interface should conform to the specification in order to function correctly. Suppose we have to maintain and understand this implementation, we need to establish the relationships between the implementation and the specification.
\begin{figure}[h!]
  \begin{center}
    \begin{tabular}{c}
\begin{minipage}[b]{\linewidth}
  \centering
  \includegraphics[width=0.8\linewidth]{code1}
\end{minipage}\\
(a) Older Version\\
\\
\begin{minipage}[b]{\linewidth}
  \centering
  \includegraphics[width=0.8\linewidth]{code2}
\end{minipage}\\
(b) Newer Version
\end{tabular}
  \caption{ Corresponding code snippets from two versions of the MAX7310 virtual device from the QEMU virtual machine. This code defines a function to reset the device.}
  %However, it is unclear why the registers are reset to the values labeled 1, 2 and 3. To know the answer, one has to refer to the specification.}
\label{fig:code}
\end{center}
\end{figure}
Figure~\ref{fig:code}(a) presents part of code extracted from the MAX7310 virtual device from QEMU. A virtual device is a software implementation emulating a hardware device. MAX7310 is a 8-bit I/O port expander. QEMU (Quick EMUlator)~\cite{Bellard05} is a virtual machine widely used in industry to assist hardware-related software development. The source code for this implementation is written in the C language. From the name of the function, we can guess it is trying to reset the device to its initial state. However, we cannot figure out what the values in the right sides of the assignment statements mean only by reading the code here. Figure~\ref{fig:spec} includes the information extracted from the specification of MAX7310. In this figure, we can see three tables for three registers: polarity register, direction register, and status register. These tables present the bit arrangement and the default value for each register. We can see that the values used in the code are defined in the specification. The code uses the hex format, and the specification uses the binary format. The piece of code and the piece of specification labeled with the same number are related, which means the piece of specification explains the piece of code with the same label.

\begin{figure}
  \centering
  \includegraphics[width=0.7\textwidth]{spec}
  \caption{Specification Example. This is part of the MAX7310 specification, including three tables which define default values for three registers.}
  \label{fig:spec}
\end{figure}

% in the directory including max7310.c: git log --follow --all -p max7310.c > a.txt
Implementation code evolves all the time in its lifecycle. By checking the QEMU repository, we found that the first version of the code in Figure~\ref{fig:code} was created on May 23, 2007, and the last change was on May 1, 2013. This code has changed 21 times till now, and it still keeps changing.
Figure~\ref{fig:code}(b) shows a newer version of the code in Figure~\ref{fig:code}(a). Line 13, 25 and 29 of the code in Figure~\ref{fig:code}(a) has changed to the line 12, 24 and 28 in Figure~\ref{fig:code}(b). Compared to code, specification is often relatively stable. The specification in Figure~\ref{fig:spec} only has three versions, and the last change happened in 2005. The specification remains unchanged since the code was created.

This example, although simple, reveals three basic requirements for managing traceability links between specification and implementations as follows.

\smallskip
\noindent 
\textbf{Requirement 1:} Traceability links should be able to uniquely identify and accurately locate relevant pieces of code in the source file and pieces of specification in the specification file. As in the example, there should be a simple and unique way to locate the code pieces and specification pieces surrounded by the boxes labeled by 1, 2 and 3.

\smallskip
\noindent
\textbf{Requirement 2:} Traceability links should be able to provide references on the statement level or expression level for source code and at the sentence level or word level for specification. As in this example, the code pieces and specification pieces surrounded by the boxes labeled by 1, 2 and 3 are all on the expression level and the word level.

\smallskip
\noindent
\textbf{Requirement 3:} Traceability links should be immune from the evolution of unrelated code. We need to ensure that an identifier created for the old version of code works for the new version if the code piece referred to by the identifier does not change.

### 2013-05-05 RPE/related.tex
revise

\section{Related Work}
\label{sec:related}
Researchers have done much work to build automatic tools for traceability link recovery (TLR) from software systems \cite{antoniol_recoveringtraceability_2002} \cite{deLucia_incremental_2006} \cite{marcus_recovering_2003}.
TLR technology scans the software artifacts including source code and documentation and recovers the semantic connections by information retrieval (IR) methods.
\todo{improve}This technology has shown to be advanced and successful \cite{spanoudakis_software_2004}.
However, IR techniques are not good enough to substitute the human decision-maker in the linking process.
Hayes et al. suggests these IR techniques should only be used to generate an appropriate list of candidate links evaluated by software analyst \cite{hayes_advancing_2006}.
The recall of these techniques is not as good as to be pratical, a significant part of traceability links need to be recovered manually.
Trace acquisition remains human-intensive and with high initial cost as reported in case studies on industrial processes and traceability experiences \cite{lindvall_practical_1996} \cite{ramesh_implementing_1995} \cite{asuncion_an_2007} \cite{gotel_extended_1997} \cite{neumuller_case_2006}.

The evolution nature of software artifacts makes the situation even worse.
No matter how the traceability links are created -- automatically or manually,
these links are probably only valid for a certain version of the software.
As software evolves during the development and maintenance process, software artifacts changes as well.
These changes might invalidate the existing traceability links.
For automatically recovered links, we can re-run the TLR tool to reestablish them.
However, this solution is computationally costly for interactive use during software development stage.
For manually recovered links, we must reestablish these links manually again, which is arduous.
This intimidate people from using TLR tools in their daily work.
Some work has been done to address this problem by improving exisiting IR based methods \cite{jiang_incremental_2008}.
However, these work didn't consider the situations specific to hardware related software.

>>>>

\section{Related Work}
\label{sec:related}
There has been much research on building automatic tools for traceability link recovery (TLR) from software artifacts \cite{antoniol_recoveringtraceability_2002} \cite{deLucia_incremental_2006} \cite{marcus_recovering_2003}.
TLR technology scans the software artifacts including source code and documentation and recovers the semantic connections by information retrieval (IR) methods.
%\todo{improve}This technology has shown to be advanced and successful \cite{spanoudakis_software_2004}.
However, IR techniques cannot completely substitute the human decision-maker in the linking process.
Hayes et al. suggests such IR techniques should only be used to generate an appropriate list of candidate links which are then evaluated by software analysts \cite{hayes_advancing_2006}.
For this reason, these automatic techniques have not been widely adopted and most traceability links still need to be recovered manually.
Link acquisition remains human-intensive and with high initial cost as reported in case studies on industrial processes and traceability experiences \cite{lindvall_practical_1996} \cite{ramesh_implementing_1995} \cite{asuncion_an_2007} \cite{gotel_extended_1997} \cite{neumuller_case_2006}.

The evolution nature of software artifacts makes the situation even worse.
No matter how the traceability links are created -- automatically or manually,
these links are probably only valid for a certain version of the software.
As software evolves during the development and maintenance process, software artifacts changes constantly.
These changes might invalidate the existing traceability links.
For manually recovered links, we must reestablish these links manually again, which is arduous.
For automatically recovered links, we can re-run the TLR tool to reestablish them.
However, this solution is computationally costly for interactive use during the software development stage.
This intimidates people from using TLR tools in their daily work.
There has been research addressing this problem by improving existing IR based methods \cite{jiang_incremental_2008}.
Our approach differs in that we try to migrate valid links instead of re-establishing these links.
%However, these work didn't consider the situations specific to hardware related software.
